---
title: "Abacus gesture study 2"
author: "Brooks et al."
date: "August 31, 2016"
output: 
  html_document:
    toc: true
    number_sections: true
    highlight: tango
    theme: spacelab
---
  
```{r include=FALSE}
require(knitr)
opts_chunk$set(
  cache=TRUE,
  warn=FALSE,
  error=FALSE,
  size="small"    # slightly smaller font for code
)
```

```{r}
library(ggplot2)
library(directlabels)
library(car)
library(lme4)
library(langcog)
library(plotrix)
library(readr)
library(knitr)
library(dplyr)
# library(plyr)

rm(list=ls())
theme_set(theme_mikabr() +
            theme(panel.grid = element_blank(),
                  strip.background = element_blank()))
font <- "Open Sans"
```

# Study 2 Experts: Graphs and analyses

Read in data (data are matlab outputs and demographics data). 

```{r}
data.bysub <- read_csv("data/study2/experts/finaldata/MAstudy2_experts_bysub.csv")
data.bytrial <- read_csv("data/study2/experts/finaldata/MAstudy2_experts_bytrial.csv")
```

Reformat factors

structure condition levels so that baseline is the reference level and the others apper in a sensible order on graphs

```{r}
data.bysub$condition <- factor(data.bysub$condition)
```

Make subnum a factor

```{r}
data.bysub$subnum <- factor(data.bysub$subnum)
```

## Graphs 

Make bar graphs showing performance on each task, with error bars
(graphs 1 and 2 not included in paper)

### Graph 1: by addend size

```{r}
threshBar <- tapply(data.bysub$max, data.bysub$condition, FUN=mean, na.rm=T)
plot <- barplot(threshBar, ylim=c(0,550), col=c("Black","dark gray","light gray","white"), main=c("Abacus Subjects: Mean Addend Size by Condition"))
ses <- tapply(data.bysub$max, data.bysub$condition, FUN=std.error)
plotCI(plot, threshBar, ses, add=T, col=c("gray"))
```

### graph 2: by threshold

```{r}
threshBar <- tapply(data.bysub$threshold, data.bysub$condition, FUN=mean, na.rm=T)
plot <- barplot(threshBar,ylim=c(0,7), ylab="Mean Threshold Level", main=c("MA Experts: Threshold Level by Condition"), col=c("Black","dark gray","light gray","white"))
ses <- tapply(data.bysub$threshold, data.bysub$condition, FUN=std.error)
plotCI(plot, threshBar, ses, add=T, col=c("gray"))
```

### Graph 3: overall threshold graph by trial

```{r}
par(cex=1.1, lwd=2)
means <- tapply(data.bytrial$threshold, list(data.bytrial$condition, data.bytrial$trial),mean,na.rm=T)
plot(means[1,1:20],type="l",col="blue", xlab=c("Trial"), ylab=c("Level"),main=c("MA Experts: Average Performance Over First 20 Trials"), lty=1)
legend("topleft",c("Baseline","Blindfold", "No Hands","Motor Interference"),lty=c(1,2,3,4), col=c("blue","green","orange","red"))
points(means[2,1:20],type="l", col="green", lty=2)
points(means[4,1:20],type="l", col="orange", lty=3)
points(means[3,1:20],type="l", col="red", lty=4)
```

## Analyses: Mixed effects model on performance. 

For analyses, consider all data with random intercept & slopes for subject

```{r}
data.bytrial$subnum <- factor(data.bytrial$subnum)

thresh.lmr <- lmer(threshold ~ condition + (condition|subnum), 
                   data = data.frame(data.bytrial))
```

And show coefficients.

```{r}
kable(summary(thresh.lmr)$coefficients, digits = 3)
```

To get simple effects p values: use t-as-z heuristic.

```{r}
 2*(1-pnorm(abs((fixef(thresh.lmr)/sqrt(diag(vcov(thresh.lmr)))))))
```

### check about other factors to possibly include in model

```{r}
lm.1 <- lmer(threshold ~ condition+ age + (1+condition|subnum) + (1|age), 
             data=data.frame(data.bytrial)) #t=.12
```

Age not significant at `r 2*(1-pnorm(abs((fixef(lm.1)/sqrt(diag(vcov(lm.1)))))))`.

```{r}
lm.1 <- lmer(threshold ~ condition+ gender + (1+condition |subnum) + (1|gender),
             data=data.frame(data.bytrial))
```

Gender not significant at `r 2*(1-pnorm(abs((fixef(lm.1)/sqrt(diag(vcov(lm.1)))))))`.

```{r}
# 
data.bytrial$order <- factor(data.bytrial$order)
lm.1 <- lmer(threshold ~ condition + order + (1+ condition|subnum) + (1|order), 
             data=data.frame(data.bytrial)) 
```

Task order not significant at `r  2*(1-pnorm(abs((fixef(lm.1)/sqrt(diag(vcov(lm.1)))))))`.

```{r}
lm.2 <- lmer(threshold ~ condition * order + (1+ condition|subnum) + (1|order), 
             data=data.frame(data.bytrial)) 
Anova(lm.2) 
```

Order * condition interaction also not significant. 


## get Z scoes on thresholds to do a population X condition interaction:

I think we want to conduct this using a single z-score for each participant for each task, rather than using every trial.

```{r}
mean <- mean(data.bysub$threshold)
data.bysub$zthresh <- scale(data.bysub$threshold)

experts.zscores <- data.bysub[,c("subnum","condition","order","zthresh", "age")]
write_csv(experts.zscores, 
          "data/study2/zscore analysis/expert_zscores.csv")
```

# US Audio Data (Novices): Graphs and analyses

Read and reformat. 

```{r}
data.bysub <- read_csv("data/study2/novices/finaldata/MAstudy2_novices_bysub.csv")
data.bytrial <- read_csv("data/study2/novices/finaldata/MAstudy2_novices_bytrial.csv")

data.bysub$condition <- factor(data.bysub$condition, 
                               levels = c("baseline", "blindfold",
                                          "stillhands","intererence"), 
                                 labels = c("Baseline", "Blindfold",
                                            "No Hands", "Motor Interference"))
data.bysub$subnum <- factor(data.bysub$subnum)
```

## GRAPH DATA (graphs 1 and 2 not included in paper)

Make bar graphs showing performance on each task, with error bars
(graphs 1 and 2 not included in paper)

### Graph 1: by addend size

```{r}
threshBar <- tapply(data.bysub$max, data.bysub$condition, FUN=mean, na.rm=T)
plot <- barplot(threshBar, ylim=c(0,25), 
                col=c("Black","dark gray","light gray","white"), 
                main=c("Control Subjects: Mean Addend Size by Condition"))
ses <- tapply(data.bysub$max, data.bysub$condition, FUN=std.error)
plotCI(plot, threshBar, ses, add=T, col=c("gray"))
```

### graph 2: by threshold

```{r}
threshBar <- tapply(data.bysub$threshold, data.bysub$condition, FUN=mean, na.rm=T)
ses <- tapply(data.bysub$threshold, data.bysub$condition, FUN=std.error)
plot <- barplot(threshBar, ylim=c(0,20), main=c("Control Participants: Threshold Level by Condition"), ylab=c("Maximum Addend Size"),col=c("Black","dark gray","light gray","white"))
plotCI(plot, threshBar, ses, add=T)
```

### Graph 3: overall threshold graph by trial

```{r}
par(cex=1.5, lwd=2)
means <- tapply(data.bytrial$threshold, list(data.bytrial $condition, data.bytrial $trial),mean,na.rm=T)
lengths <- tapply(data.bytrial $threshold, list(data.bytrial $condition, data.bytrial $trial),length)
plot(means[1,1:20],type="l",col="blue", xlab=c("Trial"), ylab=c("Level"),main=c("Control Subjects:\nAverage Performance Over 20 Trials"), lty=1, ylim=c(2,18))
legend("topleft",c("Baseline","Blindfold", "No Hands","Motor Interference"),lty=c(1,1,1,1), col=c("blue","green","orange","red"))
points(means[2,1:20],type="l", col="green", lty=2)
points(means[4,1:20],type="l", col="orange", lty=3)
points(means[3,1:20],type="l", col="red", lty=4)
```

```{r}
means <- aggregate(data.bytrial$threshold, by=list(data.bytrial$condition, data.bytrial$trial), mean, na.rm=T)
names(means) <- c("Condition","Trial","ProblemLevel")
means$Condition <- factor(means$Condition, labels=c("Baseline","Blindfold","Motor Interference","No Hands"))
means$Condition <- relevel(means$Condition, "No Hands")
means$Condition <- relevel(means$Condition, "Blindfold")
means$Condition <- relevel(means$Condition, "Baseline")

par(mfrow=c(1,2))
plot2 <- qplot(Trial, ProblemLevel, data=means, colour=Condition, xlim=c(3,20), ylim=c(2,15), main=c("Control Participants")) + geom_smooth() + theme_bw(base_size=14)+
  scale_color_brewer(palette="Set1")+
  theme(legend.justification = c(0,0), legend.position=c(0,.7))
```

## Analyses: Mixed effects model on performance. 

### Main models

For analyses, consider all data with random intercept & slopes for subject.

Note that this model doesn't converge with random slopes.


```{r}
data.bytrial$subnum <- factor(data.bytrial$subnum)
data.bytrial$condition <- factor(data.bytrial$condition)

thresh.lmr <- lmer(threshold ~ condition + (1 + condition | subnum), 
                   data = data.frame(data.bytrial))
```

And show coefficients.

```{r}
kable(summary(thresh.lmr)$coefficients, digits = 3)
```

To get simple effects p values: use t-as-z heuristic.

```{r}
 2*(1-pnorm(abs((fixef(thresh.lmr)/sqrt(diag(vcov(thresh.lmr)))))))
```

Also show for comparison the one that converges without random slopes. 

```{r}
thresh.lmr <- lmer(threshold ~ condition + (1  | subnum), 
                   data = data.frame(data.bytrial))
kable(summary(thresh.lmr)$coefficients, digits = 3)

```

### Interference as the reference level

Confusingly, this one does converge with random slopes. 

```{r}
data.bytrial$condition.i <- relevel(data.bytrial$condition, "interference")
thresh.lmr <- lmer(threshold ~ condition.i+ (1+condition.i|subnum), 
                   data=data.frame(data.bytrial))
kable(summary(thresh.lmr)$coefficients, digits = 3)
2*(1-pnorm(abs((fixef(thresh.lmr)/sqrt(diag(vcov(thresh.lmr)))))))
```

### check about other factors to possibly include in model

```{r}
lm.1 <- lmer(threshold ~ condition+ age + (1+condition|subnum) + (1|age), 
             data=data.frame(data.bytrial)) #t=.12
```

Age not significant at `r 2*(1-pnorm(abs((fixef(lm.1)/sqrt(diag(vcov(lm.1)))))))`.

```{r}
lm.1 <- lmer(threshold ~ condition+ gender + (1+condition |subnum) + (1|gender),
             data=data.frame(data.bytrial))
```

Gender not significant at `r 2*(1-pnorm(abs((fixef(lm.1)/sqrt(diag(vcov(lm.1)))))))`.

```{r}
# 
data.bytrial$order <- factor(data.bytrial$order)
lm.1 <- lmer(threshold ~ condition + order + (1+ condition|subnum) + (1|order), 
             data=data.frame(data.bytrial)) 
```

Task order not significant at `r  2*(1-pnorm(abs((fixef(lm.1)/sqrt(diag(vcov(lm.1)))))))`.

```{r}
lm.2 <- lmer(threshold ~ condition * order + (1+ condition|subnum) + (1|order), 
             data=data.frame(data.bytrial)) 
Anova(lm.2) 
```

Order * condition interaction also not significant. 

## get Z scoes on thresholds to do a population X condition interaction:

Same procedure as above.

```{r}
mean <- mean(data.bysub$threshold)
data.bysub$zthresh <- scale(data.bysub$threshold)

novices.zscores <- data.bysub[,c("subnum","condition","order","zthresh", "age")]
write_csv(novices.zscores, 
          "data/study2/zscore analysis/novice_zscores.csv")
```

# Z-score analysis

```{r}
novices.zscores$pop <- "novices"
experts.zscores$pop <- "experts"
all <- bind_rows(novices.zscores, experts.zscores)
summary(all) #need to make subnum and order factors
all$subnum <- factor(all$subnum)
all$order <- factor(all$order)
all$pop <- factor(all$pop)
```

```{r}
z.lm <- lmer(zthresh ~ condition * pop + (1 |subnum), data=all)
kable(summary(z.lm)$coefficients, digits = 3)
2*(1-pnorm(abs((fixef(z.lm)/sqrt(diag(vcov(z.lm))))))) #interaction significant
```


## Only including experts over age 11

```{r}
sort(experts.zscores$age)
older <- experts.zscores[experts.zscores$age > 11,]
sort(novices.zscores$age)
```

```{r}
all.old <- rbind(novices.zscores, older)
summary(all.old)
all.old$subnum <- factor(all.old$subnum)
all.old$order <- factor(all.old$order)
all.old$pop <- factor(all.old$pop)
```

```{r}
z.lm <- lmer(zthresh ~ condition * pop + (1 |subnum), data=all.old)
kable(summary(z.lm)$coefficients, digits = 3)
2*(1-pnorm(abs((fixef(z.lm)/sqrt(diag(vcov(z.lm))))))) #interaction significant
```

## Only including experts over age 12

```{r}
sort(experts.zscores$age)
older <- experts.zscores[experts.zscores$age > 12,]
sort(novices.zscores$age)
all.old <- rbind(novices.zscores, older)
summary(all.old)
all.old$subnum <- factor(all.old$subnum)
all.old$order <- factor(all.old$order)
all.old$pop <- factor(all.old$pop)
z.lm <- lmer(zthresh ~ condition * pop + (1 |subnum), data=all.old)
kable(summary(z.lm)$coefficients, digits = 3)
2*(1-pnorm(abs((fixef(z.lm)/sqrt(diag(vcov(z.lm))))))) #interaction significant
```