---
title: "Abacus gesture study 1"
author: "Brooks et al."
date: "August 31, 2016"
output: 
  html_document:
    toc: true
    number_sections: true
    highlight: tango
    theme: spacelab
---

```{r include=FALSE}
require(knitr)
opts_chunk$set(
  cache=TRUE,
  warn=FALSE,
  error=FALSE,
  size="small"    # slightly smaller font for code
)
```

```{r}
library(ggplot2)
library(directlabels)
library(car)
library(lme4)
library(langcog)
library(dplyr)
library(knitr)
# library(plyr)

rm(list=ls())
theme_set(theme_mikabr() +
            theme(panel.grid = element_blank(),
                  strip.background = element_blank()))
font <- "Open Sans"
```

# Read in data

```{r}
data = read.csv("data/study1/Combined 2011 2012 coded data 11-25-13.csv")
```

Data cleaning.

+ round difficulty ratings so they make up distinct groups
+ create a binomial measure of gesture size
+ create a binomial measures of gesture present
+ exclude participant 2455 - only 5 years old

```{r}
data$rdif=round(data$difficulty)
data$size.01 <- ifelse(data$size > 2, 1, ifelse(data$size < 3, 0, NA))
data$g.present = ifelse(data$g.present==9, NA, data$g.present)
data = data[data$subnum!=2455,]
```

# Descriptives

Proportion of trials containing gesture is `r mean(data$g.present, na.rm=T)`.

Now look at the the trials with no gesture.

```{r}
length(data[data$g.present==0 & !is.na(data$g.present),]$g.present) #168 trials
nog <- data[data$g.present==0 & !is.na(data$g.present),]
hist(nog$level)
sort(nog$level)
```

129 (77%) were on trials of level 3 or lower, meaning that there were only 2 addends in the problem. 

# Statistical models

consider the following variables: level, difficulty, # of moves,trial

```{r}
hist(data$level)
hist(data$difficulty)
hist(data$moves)
hist(data$trial)
```

We ran a mixed effects regression model predicting size as a continuous variable with Subjective Difficulty, Trial Number, log of Problem Level (our first measure of objective difficulty), and Number of Steps Required (our second measure of objective difficulty) as independent variables, with a random intercept of subject and random slopes for all four independent variables. 

```{r}
sizefull <- lmer(size ~ difficulty + moves + trial + log(level) + 
                   (difficulty + moves + trial + log(level)| subnum), 
                 data=data) 
```

And show coefficients in a nice table. 

```{r}
kable(summary(sizefull)$coefficients, digits = 3)
```

Binomial analyses

```{r}
size.allfour.binom <- glmer(size.01 ~ difficulty + moves + trial + log(level) + 
                              (difficulty + moves + trial + log(level)| subnum), 
                            data = data, family="binomial")
```

And show coefficients.

```{r}
kable(summary(size.allfour.binom)$coefficients, digits = 3)
```

# Plot

Size by level and difficulty graph

get median difficulty level by problem level

```{r}
ms <- aggregate(difficulty ~ level + subnum, data=data, mean)
meds <- aggregate(difficulty ~ level, data=ms, median)
names(meds) <- c("level","med")
data2 <- merge(data, meds, by="level", all.x=T)
data <- data2
data$Difficulty_Category <- ifelse(data$difficulty < data$med, 
                                   "Below Median Difficulty (Easier)", 
                                   "At or Above Median Difficulty (Harder)")
```

More aggregation.

```{r}
data$Subjective_Difficulty <- data$Difficulty_Category
ms1 <- aggregate(size ~ level + Subjective_Difficulty + subnum, data=data, mean)
ms2 <- aggregate(size ~ level + Subjective_Difficulty, data=ms1, mean)
data$Subjective_Difficulty <- factor(data$Subjective_Difficulty)
ms2$measure.type <- "Gesture Size"
ms2$measure <- ms2$size
ms <- bind_rows(ms2)
```

And plot.

```{r}
qplot(level,measure, colour=Subjective_Difficulty,
      label=c(Subjective_Difficulty),
      data=ms, 
      ylab=c("Average Gesture Size"), 
      xlab=c("Problem Level (Number of Addends - 1)"), ylim=c(1,4)) +
  geom_smooth() +
  theme_bw(base_size=14) +
  scale_color_brewer(palette="Set1")+
  theme(legend.justification = c(0,0), legend.position=c(.6,.7))
```

# Control for coder

Try this out. 

```{r}
sizefull <- lmer(size ~ difficulty + moves + trial + log(level) + 
                   (difficulty + moves + trial + log(level)| subnum), 
                 data=data) 
```

And show coefficients in a nice table. 

```{r}
kable(summary(sizefull)$coefficients, digits = 3)
```